Local simple connectedness of boundaries of hyperbolic groups

### 2020-03-26 paper.tex
Proof read 4.1

We begin this section by recalling work of Bestvina and Mess, Bowditch, and
Swarup, which characterise the hyperbolic groups with locally path connected
boundary. In~\cite{bestvinamess91}, Bestvina and Mess introduce the following
condition that a hyperbolic group might satisfy.

@@

\begin{proposition}\cite[Proposition 3.2]{bestvinamess91}
  If $G$ satisfies $\ddag_{11\delta + 3}$ then $\boundary G$ is locally path
  connected.
\end{proposition}

\begin{proposition}\cite[Proposition ?]{bestvinamess91}\label{prop:BM2}
  If $\boundary G - \set\xi$ is connected for any $\xi \in \boundary G$ then
  $G$ satisfies $\ddag_M$ for any $M$.
\end{proposition}

@@

In this section we introduce a related condition. This condition is stronger
than $\ddag_M$, but, using Bestvina and Mess's result, we show that it holds
under the same assumption. This condition reinterprets the paths in $\ddag_M$
as paths in the complexes $K_n$, with the additional assumption that these
paths approximate paths that factor through the projection map $p^\infty_n$.
This interpretation gives a more systematic method for constructing paths in
$\boundary G$ from paths in the spheres $K_n$, and permits a generalisation to
higher dimensional connectivity.

\begin{definition}
  We say that $G$ satisfies $\ddag'_M$ if there exist numbers $L$ and $n_0$ such that,
  for any $n \geq n_0$, any $\zeta_1$ and $\zeta_2$ in $\boundary G$
  such that $\newdist(p^\infty_n(\zeta_1), p^\infty_n(\zeta_2)) \leq M$, and
  any vertices $x_1$ and $x_2$ of $K_n$ adjacent to $p^\infty_n(\zeta_1)$ and
  $p^\infty_n(\zeta_2)$, there is a path $\alpha\from I\to \boundary G$ from
  $\zeta_1$ to $\zeta_2$ in $\boundary G$ such that $p^\infty_n\composed\alpha$
  admits a simplicial approximation connecting $x_1$ to $x_2$ in $K_n$ after at
  most $L$ barycentric subdivisions of $I$.
\end{definition}

@@

\begin{proof}
  Suppose $\ddag'_M$ does not hold. Then for any $i$ there exists a number $n_i
  \geq i$, points $\zeta_1^i$ and $\zeta_2^i$ in $\boundary G$ and vertices
  $x_1^i$ and $y_1^i$ of $K_{n_i}$ adjacent in $K_{n_i}$ to
  $p^\infty_{n_i}(\zeta^i_1)$ and $p^\infty_{n_i}(\zeta^i_2)$ such that there
  is no path $\alpha$ from $\zeta_1^i$ to $\zeta_2^i$ in $\boundary G$ where
  $p^\infty_{n_i}\composed\alpha$ admits a simplicial approximation after at
  most $i$ barycentric subdivisions of the interval.

  Take group elements $g_i$ such that $g_i\cdot x_1^i$ is equal to some vertex
  $x_1$ for all $i$ vertices and pass to a subsequence as in
  Section~\ref{sec:horoball}, so that $g_i\to\xi$ as $i \to\infty$. Since
  $\dist(x_1^i, x_2^i)$ is bounded we can pass to a deeper subsequence so that
  $g_i\cdot x_2^i$ is constant equal to $x_2$ for all $i$, and $x_2$ is
  contained in $H_i$ for all $i$.

  Now, $g_i\cdot \zeta_1^i$ and $g_i\cdot\zeta_2^i$ map by $q_i$ to points in
  simplices adjacent to $x_1$ and $x_2$, so by Lemma~\ref{lem:propermaps} we
  can pass to a subsequence so that the sequences $g_i\cdot \zeta_1^i$ and
  $g_i\cdot\zeta_2^i$ converge to points $\zeta_1$ and $\zeta_2$ in $\boundary
  G - \set\xi$ as $i \to \infty$.

  Let $C$ be a compact subset of $\boundary G - \set\xi$ containing a path
  $\beta$ from $\zeta_1$ to $\zeta_2$ and also a neighbourhood of each of the
  points $\zeta_1$ and $\zeta_2$. By Lemma~\ref{lem:uniform_continuity} and
  compactness of $C$, there exists $\epsilon > 0$ and $i_0 \in \naturals$ such
  that for subset $V$ of $C$ of diameter at most $\epsilon$ and any $i \geq
  i_0$, $q_i(V)$ has diameter (with respect to $\newdist$) at most $6\delta+2$.
  By increasing $i_0$ we may also assume that $\zeta_1^i$ and $\zeta_2^i$ are
  contained in $C$ for $i \geq i_0$. By decreasing $\epsilon$ we may assume
  that the $\epsilon$-neighbourhoods of $\zeta_1$ and $\zeta_2$ are contained
  in $C$.

@@

\begin{proof}
  For the first claim, see~\cite{bestvinamess91}. For the second, let $p$ be a 
  point on $\alpha$ such that $\dist(x, p) \leq \delta$. Then the triangle 
  inequality shows that $\abs{\dist(e,p) - \dist(e,x)} \leq \delta$, and 
  so $\dist(p, \alpha(\dist(e,x))) \leq \delta$. Then the result 
  follows from another application of the triangle inequality.
\end{proof}

\begin{proposition}\label{prop:i_in_dimension_1}
  Suppose that $G$ satisfies $\ddag'_{D + 4\delta+2}$. Then there exist $L$ and
  $n_0$ such that for $n\geq n_0$ there is a map $i^n_{n+1}\from \Skel_1 K_n
  \to \Skel_1 K_{n+1}$ satisfying the following conditions:
  \begin{enumerate}
    \item for each $n$, $i^n_{n+1}$ is an affine map on the subdivision
      $\sd_L\Skel_1 K_n$,
    \item for any vertex $x$ of $K_n$, $\dist(x, i^n_{n+1}(x)) \leq 2\delta+1$,
      and
    \item for each $n$, the map $i^n_{n+1}$ is a simplicial approximation a map
      that factors through $p^\infty_{n+1}\from \boundary G \to K_{n+1}$.
  \end{enumerate}
\end{proposition}

@@

  For an edge $e$ of $K_n$ connecting vertices $x_1$ and $x_2$,
  $\dist(i^n_{n+1}(x_1), i^n_{n+1}(x_2))$ is at most $D + 4\delta + 2$. It
  follows that there is a path $\alpha$ from $\gamma_{x_1}(\infty)$ to
  $\gamma_{x_2}(\infty)$ in $\boundary G$ so that
  $p^\infty_{n+1}\composed\alpha$ admits a simplicial approximation connecting
  $i^n_{n+1}(x_1)$ to $i^n_{n+1}(x_2)$ after at most $L$ subdivisions. Define
  $i^n_{n+1}$ to be this simplicial map on $\sd^L e$.

>>>>

We begin this section by recalling work of Bestvina and Mess, Bowditch, and
Swarup, which characterise the hyperbolic groups with locally path connected
boundary. In~\cite{bestvinamess91}, Bestvina and Mess introduce the following
condition that a hyperbolic group might satisfy. (Regarding the constants,
note that with our convention that ideal triangles are $\delta$-hyperbolic,
$C$ in the cited paper can be taken to be equal to $\delta$.)

@@

\begin{proposition}\cite[Proposition 3.2]{bestvinamess91}
  If $G$ satisfies $\ddag_{8\delta + 3}$ then $\boundary G$ is locally path
  connected.
\end{proposition}

\begin{proposition}\cite[Proposition 3.3]{bestvinamess91}\label{prop:BM2}
  If $\boundary G - \set\xi$ is connected for all $\xi \in \boundary G$ then
  $G$ satisfies $\ddag_M$ for any $M$.
\end{proposition}

@@

In this section we introduce a related condition. This condition is stronger
than $\ddag_M$, but, using Bestvina and Mess's result, we show that it holds
under the same assumption. This condition reinterprets the paths in $\ddag_M$
as edge paths in the complexes $K_n$, with the additional assumption that
these paths approximate paths that factor through the projection map
$p^\infty_n$. This interpretation strengthens the link between
path-connectedness of the boundary and connectedness of the complexes $K_n$.

\begin{definition}
  We say that $G$ satisfies $\ddag'_M$ if there exist numbers $L$ and $n_0$
  such that, for any $n \geq n_0$, any $\zeta_1$ and $\zeta_2$ in $\boundary
  G$ such that $\newdist(p^\infty_n(\zeta_1), p^\infty_n(\zeta_2)) \leq M$,
  and any vertices $x_1$ and $x_2$ of the minimal simplices of $K_n$
  containing $p^\infty_n(\zeta_1)$ and $p^\infty_n(\zeta_2)$, there is a path
  $\alpha\from I\to \boundary G$ from $\zeta_1$ to $\zeta_2$ in $\boundary G$
  such that $p^\infty_n\composed\alpha$ admits a simplicial approximation
  connecting $x_1$ to $x_2$ in $K_n$ after at most $L$ barycentric
  subdivisions of $I$.
\end{definition}

@@

\begin{proof}
  Suppose $\ddag'_M$ does not hold. Then for any $i$ there exists a number $n_i
  \geq i$, points $\zeta_1^i$ and $\zeta_2^i$ in $\boundary G$ and vertices
  $x_1^i$ and $y_1^i$ of the minimal simplex of $K_{n_i}$ containing 
  $p^\infty_{n_i}(\zeta^i_1)$ and $p^\infty_{n_i}(\zeta^i_2)$ such that there
  is no path $\alpha$ from $\zeta_1^i$ to $\zeta_2^i$ in $\boundary G$ where
  $p^\infty_{n_i}\composed\alpha$ admits a simplicial approximation after at
  most $i$ barycentric subdivisions of the interval.

  Take group elements $g_i$ such that $g_i\cdot x_1^i$ is equal to some vertex
  $x_1$ for all $i$ vertices and pass to a subsequence as in
  Section~\ref{sec:horoball}, so that $g_i\to\xi$ as $i \to\infty$. Since
  $\dist(x_1^i, x_2^i)$ is bounded we can pass to a deeper subsequence so that
  $g_i\cdot x_2^i$ is constant equal to some vertex $x_2$ for all $i$, and
  $x_2$ is contained in $H_i$ for all $i$.

  Now, $g_i\cdot \zeta_1^i$ and $g_i\cdot\zeta_2^i$ map by $q_i$ to points in
  simplices containing $x_1$ and $x_2$, so by Lemma~\ref{lem:propermaps} we
  can pass to a subsequence so that the sequences $g_i\cdot \zeta_1^i$ and
  $g_i\cdot\zeta_2^i$ converge to points $\zeta_1$ and $\zeta_2$ in $\boundary
  G - \set\xi$ as $i \to \infty$.

  Let $C$ be a compact subset of $\boundary G - \set\xi$ containing a path
  $\beta$ from $\zeta_1$ to $\zeta_2$ and also a neighbourhood of each of the
  points $\zeta_1$ and $\zeta_2$. By Lemma~\ref{lem:uniform_continuity}, there
  exists $\epsilon > 0$ and $i_0 \in \naturals$ such that for subset $V$ of
  $C$ of diameter at most $\epsilon$ and any $i \geq i_0$, $q_i(V)$ has
  diameter (with respect to $\newdist$) at most $6\delta+2$. By increasing
  $i_0$ we may also assume that $\zeta_1^i$ and $\zeta_2^i$ are contained in
  $C$ for $i \geq i_0$. By decreasing $\epsilon$ we may assume that the
  $\epsilon$-neighbourhoods of $\zeta_1$ and $\zeta_2$ are contained in $C$.

@@

\begin{proof}
  For the first claim, see~\cite{bestvinamess91} and recall that ideal
  triangles are assumed to be $\delta$-hyperbolic. For the second, let $p$ be
  a point on $\alpha$ such that $\dist(x, p) \leq \delta$. Then the triangle
  inequality shows that $\abs{\dist(e,p) - \dist(e,x)} \leq \delta$, and so
  $\dist(p, \alpha(\dist(e,x))) \leq \delta$. Then the result follows from
  another application of the triangle inequality.
\end{proof}

\begin{proposition}\label{prop:i_in_dimension_1}
  Suppose that $G$ satisfies $\ddag'_{D + 4\delta+2}$. Then there exist $L$ and
  $n_0$ such that for $n\geq n_0$ there is a map $i^n_{n+1}\from \Skel_1 K_n
  \to \Skel_1 K_{n+1}$ satisfying the following conditions:
  \begin{enumerate}
    \item for each $n$, $i^n_{n+1}$ is a simplicial map on the subdivision
      $\sd_L\Skel_1 K_n$,
    \item for any vertex $x$ of $K_n$, $\dist(x, i^n_{n+1}(x)) \leq 2\delta+1$,
      and
    \item for each $n$, the map $i^n_{n+1}$ is a simplicial approximation a map
      that factors through $p^\infty_{n+1}\from \boundary G \to K_{n+1}$.
  \end{enumerate}
\end{proposition}

@@

  For an edge $e$ of $K_n$ connecting vertices $x_1$ and $x_2$,
  $\dist(i^n_{n+1}(x_1), i^n_{n+1}(x_2))$ is at most $2(2\delta + 1) + D = D +
  4\delta + 2$. It follows from $\ddag'_{D+4\delta+2}$ that there is a path
  $\alpha$ from $\gamma_{x_1}(\infty)$ to $\gamma_{x_2}(\infty)$ in $\boundary
  G$ so that $p^\infty_{n+1}\composed\alpha$ admits a simplicial approximation
  connecting $i^n_{n+1}(x_1)$ to $i^n_{n+1}(x_2)$ after $L$ subdivisions.
  Define $i^n_{n+1}$ to be this simplicial map on $\sd^L e$.
